---
title: "STAT 5014 Homework 5"
author: "Chaoran Wang"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  word_document: default
header-includes: \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stargazer)
library(swirl)
library(data.table)
library(svTools)
library(vioplot)
library(xlsx)
library(ggplot2)
library(kableExtra)
library(ggExtra)
library(doBy)
library(GGally)
library(ellipse)
library(psych)
library(fiftystater)
```

## Problem 3
A good figure should shows clear information to readers in a second which means readers should derive the information we want them know quickly. It should includes clear title, axis labels, necessary values, and possible highlights. The plot should comes with appropriate type of graphs. For example, histogram is better to show the distribution of the data. If you want to focus on the relationship, especially the regression, scatter plot with a fitted line might be better. Box plot is better to show the basic statistics of the data like mean and medium. If you want to focus on the proportion of a set of categorical data with few categories, pie chart should be the best one. If it has many categories, bar chart should be better.

## Problem 4
I first create the function to compute the proportion of successes in a vector. Then I use 'apply' function to compute the proportion of success in P4b_data by row and column. we can see the two results are different. The rows coming with all "1" always have 100% of success when we 'apply' by row. There is no column coming wih all "1", so the results are some proportions when we 'apply' by column. Therefore, using 'apply' by column seems make more sense. **However, both row and column seems to be incorrect because the codes fill the matrix just with the repeats of the first row/column it get, so we always get same result for each row/column which is not random.** Codes are in Appendix.

```{r Problem4_fun, echo=F, eval=T}
########################### 
#Problem4_fun
#function to compute the proportion of successes in a vector
###########################
compute_prop <- function(x){
  n <- length(x)
  num_S <- sum(x == 1) 
  prop_S <- num_S/n
}
```

```{r Problem4_simulation, echo=F, eval=T}
########################### 
#Problem4_simulation
###########################
P4b_data <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10, ncol = 10)
prob_row <- apply(P4b_data, 1, function(x) compute_prop(x))
prob_row
prob_col <- apply(P4b_data, 2, function(x) compute_prop(x))
prob_col
```

Next, I create a function to simulate 10 flips of a coin when a probability is given. Then, I assign a vector of the desired probabilities to be (10, 20, ..., 90). Using the simulation function, I get a matrix back. Using the 'Problem4_fun', I can prove the simulation function does work.
```{r Problem4_fun_prob, echo=F, eval=T}
########################### 
#Problem4_fun_prob
#function to simulate 10 flips of a coin given a probability
###########################
simulate_prob <- function(x){
  simulate_data <- rbinom(10, 1, prob = x/100)
}
```

```{r Problem4_simu_prob, echo=F, eval=T}
########################### 
#Problem4_simu_prob
#function to simulate 10 flips of a coin given a probability
###########################
prob <- seq(10, 90, 10)
simu_res <- sapply(prob, function(x) simulate_prob(x))
res_proof <- apply(simu_res, 2, function(x) compute_prop(x))
knitr::kable(rbind(prob,res_proof,simu_res), format = "latex", booktabs=T, 
             caption="Simulation when Probabilities are Given") %>% 
  kable_styling(latex_options = "hold_position") %>%
  group_rows("Simulation", 2, 3, latex_gap_space = "2em")
```

\newpage
## Problem 5
All codes of this problem are in Appendix. 

I first load and tidy the data, then print a basic summary table out by starch groups.
```{r Problem5_starch_analysis, echo=F, eval=T}
###########################
#Problem5_starch_analysis  
#get data  
###########################
url <- "http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat"
starch_raw <- read.table(url, header = F, skip = 1, fill = T, stringsAsFactors = F)
colnames(starch_raw) <- c("starch", "strength", "thickness")
starch_sum <- describeBy(starch_raw[,2:3], starch_raw$starch, mat=TRUE)
knitr::kable(starch_sum[,c(2,4:7,10:13)],caption="Summary by starch Group")
```

Then, I explore the data and print necessary figures out. All figures are based on starch groups.
```{r Problem5_multiplot, echo=F, eval=T}
###########################
# Problem5_multiplot
# Multiple plot function
# Credit to Cookbook for R
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
###########################
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r Problem5_plot, echo=F, eval=T, include=T, fig.height=8, fig.width=9}
########################### 
#Problem5_plot
###########################
p1 <- ggplot(starch_raw, aes(x=thickness, y=strength, colour=starch, group=starch)) +
   geom_point() + 
   ggtitle("Scatter Plot for strength vs. thickness")
p2 <- ggplot(starch_raw,  aes(x=thickness, y=strength, colour=starch, group=starch)) +
   geom_point() + stat_smooth(method = "lm") +
   ggtitle("Regression of strenght on thickness")
p3 <- ggplot(starch_raw, aes(x=thickness, colour=starch)) +
   geom_density() +
   ggtitle("Density of thickness by starch")
p4 <- ggplot(starch_raw, aes(x=strength, colour=starch)) +
   geom_density() +
   ggtitle("Density of strength by starch")
p5 <- ggplot(starch_raw, aes(x=thickness, y=strength)) +
     geom_point() + facet_wrap(~starch) +
     ggtitle("Scatter Plot for strength vs. thickness by starch")
p6 <- ggplot(starch_raw, aes(x=strength, fill=starch)) +
   geom_histogram(colour="black", binwidth=30) +
   ggtitle("Histogram of strength by starch")
p7 <- ggplot(starch_raw, aes(x=starch, y=thickness)) +
   geom_boxplot(stat="boxplot", aes(colour = starch)) +
   ggtitle("Boxplot of thickness by starch")
p8 <- ggplot(starch_raw, aes(x=starch, y=strength)) +
   geom_boxplot(stat="boxplot", aes(colour = starch)) +
   ggtitle("Boxplot of strength by starch")
multiplot(p1, p2, p3, p4, p5, p6, p7, p8, cols=2)
```


```{r Problem5_plot_mar, echo=F, eval=T, include=T, fig.height=8, fig.width=8}
########################### 
#Problem5_plot_mar
###########################
p <- ggplot(starch_raw, aes(thickness, strength)) + geom_point() + theme_classic() +
  ggtitle("Density of a Scatter Marginal Plot")
ggMarginal(p, starch_raw, type = "histogram", yparams=list(colour="orange"))
```

\newpage
Finally, I print the plot to show the correlations between variables.

```{r Problem5_plot_cor, echo=F, eval=T, include=T, fig.height=8, fig.width=9}
########################### 
#Problem5_plot_cor
###########################
# Basic Scatterplot Matrix
ggpairs(starch_raw, aes(colour = starch, alpha = 0.4), title = "Simple Scatterplot Matrix", 
        lower = list(combo = wrap("facethist", binwidth = 0.5)))
ggplot(starch_raw, aes(x=thickness, y=strength, colour=starch)) + geom_point() + 
   stat_ellipse() + ggtitle("Scatter Plot and Correlations for strength vs. thickness")
```

## Problem 6

All codes used in this problem are in Appendix.
```{r problem6_data, echo=F, eval=T}
########################### 
#problem6_data
###########################
# we are grabbing a SQL set from here
# http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

# download the files, looks like it is a .zip
# library(downloader)
# download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",
# dest = "us_cities_states.zip")
# unzip("us_cities_states.zip", exdir = "D:/STAT_5014/05_R_apply_family")

# read in data, looks like sql dump, blah
library(data.table)
states <- fread(input = "./us_cities_and_states/states.sql",
 sep = "'", sep2 = ",", header = F, select = c(2,
4))
colnames(states) <- c("State", "Abbr")
cities <- fread(input = "./us_cities_and_states/cities_extended.sql",
 sep = "'", sep2 = ",", header = F, select = c(2,
4))
colnames(cities) <- c("Cities", "State")
### YOU do the CITIES I suggest the cities_extended.sql
### may have everything you need can you figure out how to
### limit this to the 50?
# delete DC
states_50 <- states[-8,]
states_50$State <- tolower(states_50$State)
# delete PR & DC
cities_50 <- cities[which(State != "PR" & State != "DC"),]
```

```{r problem6_sum, echo=F, eval=T}
count_cities <- data.frame(table(cities_50$State))
count_cities <- cbind(states_50$State,count_cities)
colnames(count_cities) <- c("State", "Abbr", "Cities_Count")
knitr::kable(count_cities[1:5,2:3],caption="The Number of Cities by State (first 5 rows)")
knitr::kable(summary(count_cities[,2:3]),caption="Summary of the Number of Cities by State")
```

```{r problem6_letter_count, echo=F, eval=T}
########################### 
#problem6_letter_count
###########################
##pseudo code
letter_count <- data.frame(matrix(NA,nrow=50, ncol=26))
colnames(letter_count) <- c(letters)
rownames(letter_count) <- c(states_50$Abbr)
getCount <- function(letter,state_name){
    temp <- strsplit(state_name,"")
    count_letter <- data.frame(matrix(NA, nrow=1, ncol=26))
    colnames(count_letter) <- c(letters)
    for (i in 1:26){
      count_letter[,i]<-sapply(letter[i], function(x) x<-sum(x==unlist(temp)))
    }
    return(count_letter)
}

for(j in 1:50){
    letter_count[j,] <- getCount(letters,states_50[j,State])
}
```

```{r problem6_maps, echo=F, eval=T, include=T, fig.height=6, fig.width=8}
########################### 
#problem6_maps
###########################
##pseudo code
# https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html
data("fifty_states") # this line is optional due to lazy data loading
# map_id creates the aesthetic mapping to the state name
# column in your data
p1 <- ggplot(count_cities, aes(map_id = State)) + # map points to the fifty_states shape data
geom_map(aes(fill = Cities_Count), map = fifty_states) + expand_limits(x = fifty_states$long,
y = fifty_states$lat) + coord_map() + scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) + labs(x = "", y = "") +
  theme(legend.position = "bottom", panel.background = element_blank()) +
  ggtitle("Map 1: Colored by Count of Cities within States")

p1

letter_count["lgt3"] <- ifelse(apply(letter_count, 1, function(x) any(x > 3)),"TRUE", "FALSE")
letter_count_states <- cbind(states_50$State,letter_count)
colnames(letter_count_states)[1] <- "State"

p2 <- ggplot(letter_count_states, aes(map_id = State)) + # map points to the fifty_states shape data
geom_map(aes(fill=lgt3), map = fifty_states) + expand_limits(x = fifty_states$long,
y = fifty_states$lat) + coord_map() + scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) + labs(x = "", y = "") +
  theme(legend.position = "bottom", panel.background = element_blank()) +
  ggtitle("Map 2: Highlights of States with 3 occurances of ANY letter")

p2

# ggsave(plot = p, file =
# 'HW5_Problem6_Plot_Settlage.pdf')
```

\newpage
#Appendix 1: R code  
```{r Appendix, ref.label=c("Problem4_fun", "Problem4_simulation", "Problem4_fun_prob", "Problem4_simu_prob", "Problem5_starch_analysis", "Problem5_multiplot", "Problem5_plot", "Problem5_plot_mar", "Problem5_plot_cor", "problem6_data", "problem6_letter_count", "problem6_maps"), echo=TRUE, eval=F, tidy=FALSE, tidy.opts=list(blank=FALSE, width.cutoff=20), include=T}
  
```