---
title: "STAT 5014 Homework 3"
author: "Chaoran Wang"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  word_document: default
header-includes: \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stargazer)
library(swirl)
library(data.table)
library(svTools)
library(vioplot)
```

## Fineshed Problem 1 to 2

## Problem 4
What I takeaway from these guides is it is very important to learn a good programming style which makes my codes be easy to read, share, and verify. For example, I used "=" more than "<-" before which seems not to be good. And sometimes, I liked to named some regression to be lm1 which is not clear for an object name. Besides, I hope I could be more familiar with "%>%" operator in dplyr package later. At this time, the statement in order seems to be more familiar to me. I am always confused about how to use "%>%" when coding.

## Problem 5
I cannot find syntax error using 'lint' function.

```{r Problem5_lintr, echo=F, eval=T}
########################### 
#Problem5_lintr  
#get data  
###########################
lint(file = "D:/STAT_5014/02_data_munging_summarizing_R_git/HW2_Wang_Chaoran.Rmd")
##########################
```

## Problem 6
I first create the function which calculate the statistics we need. Then I read in the data and tidy it. Next, I apply the function to the data by group as Observer and print the results we need. The codes are in Appendix.

```{r Problem6_func, eval=T, echo=F, cache=F, include=T}
########################### 
#Problem6_func
###########################
# Function to create a summary statistic
sum_stat <- function(x){
  dev1 <- x[,1]
  dev2 <- x[,2]
  mean_dev1 <- mean(dev1)
  mean_dev2 <- mean(dev2)
  sd_dev1 <- sd(dev1)
  sd_dev2 <- sd(dev2)
  cor_dev1dev2 <- cor(dev1, dev2)
  result <- data.frame(mean_dev1, mean_dev2, sd_dev1, sd_dev2, cor_dev1dev2)
  return(result)
}
###########################
```

When I try to plot the boxplots and violin plots of dev1 and dev2 together, the plots just show as a straight line. I guess that is because of the ranges of each column are small. The means of dev1 are all about 54.26 while the means of dev2 are all about 47.83. That makes the two ranges lie far away from each other and look like a straight line. The same thing happens to violin plot. Hence, I decide to split them into different plots which look better.

```{r Problem6_data, eval=T, echo=F, cache=F, include=T, results='asis', fig.width=7, fig.height=4}
########################### 
#Problem6_data
###########################
# Read data and tidy data
data_pro6 <-  readRDS(file = "D:/STAT_5014/03_good_programming_R_functions/HW3_data.rds") %>%
  arrange(Observer)
# Collecting the summary statistics via sum_stat function 
# Using by function here which is easiser than looping
sum_list <- by(data_pro6[, 2:3], data_pro6[,1], sum_stat)
# Convert the list to data.frame
sum_result <- data.frame(matrix(unlist(sum_list), nrow=13, byrow=T))
Observer <- c(1:13)
sum_result <- cbind(Observer, sum_result)
colnames(sum_result) <- c("Observer", "mean_dev1", "mean_dev2", "sd_dev1", "sd_dev2", "cor_dev1dev2")
knitr::kable(sum_result,caption="Summary Statistics of 13 Observers")
par(mfrow=c(1,2))
boxplot(sum_result[,2:3])
title("Box plot of means")
vioplot(sum_result[,4],sum_result[,5],names=c("sd_dev1", "sd_dev2"), col="gold")
title("Violin plot of sds")
par(mfrow=c(1,2))
boxplot(sum_result[,2])
boxplot(sum_result[,3])
par(mfrow=c(1,2))
vioplot(sum_result[,4],names="sd_dev1", col="gold")
vioplot(sum_result[,5],names="sd_dev2", col="gold")
##########################
```

## Problem 7
First, I read in and create a tidy dataset. To tidy the dataset, I first removed the reduplicated "Day" columns and the first row with names. Then I rename the columns. I then using gather() function to tidy all readings by devices or doctors. Then I assign 1 to 6 to three devices and three doctors to be indicators of different reading methods.

After tidying, a summary is in Table 1. Since I am not sure what the experiment is actually, I cannot analyze the data based on the information I have. It looks not like a simple linear model. The codes are in Appendix.
```{r Problem7_BP_analysis, echo=F, eval=T}
########################### 
#Problem7_BP_analysis  
#get data  
###########################
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BloodPressure.dat"
BP_raw <- read.table(url, header = F, skip = 1, fill = T, stringsAsFactors = F)
BP_tidy <- BP_raw[-1,-5]
colnames(BP_tidy) <- c("Day", "Device_1", "Device_2", "Device_3", "Doctor_1", "Doctor_2", "Doctor_3")
BP_tidy$Day <- as.numeric(as.character(BP_tidy$Day))
BP_tidy <- gather(BP_tidy, Method, value, Device_1:Doctor_3) %>%
  separate(Method, c("Method", "ID"), "_") %>%
  arrange(Day)
BP_tidy$value <- as.numeric(as.character(BP_tidy$value))
BP_tidy$ID <- as.numeric(as.character(BP_tidy$ID))
###########################
```

```{r Problem7_BP_fig, eval=T, echo=F, cache=F, include=T, results='asis', fig.cap="Blood Pressure Reading by Devices and Doctors",fig.width=3, fig.height=3}
########################### 
#Problem7_BP_analysis  
#plot 
###########################
knitr::kable(summary(BP_tidy),caption="Blood Pressure Data Summary")
###########################
```

## Problem 8
I first create a function based on my math knowledge about Newton's Method. The codes are in Appendix.
```{r Problem8_func, eval=T, echo=F, cache=F, include=T}
########################### 
#Problem8_func
###########################
# Function of Newton's Method
newton <- function(fun, tol=1e-20, x0, N) {
  h <- 0.001
  i <- 1
  x1 <- x0
  z <- numeric(N)
  while (i <= N) {
    df_dx <- (fun(x0 + h) - fun(x0)) / h
    error <- fun(x0)/df_dx
    x1 <- (x0 - error)
    z[i] <- x1
    i <- i + 1
    if (abs(x1 - x0) < tol) 
    break
    x0 <- x1
    }
  return(z[1:(i-1)])
}
###########################
```

Since the function has sin and cos part, the numbers of roots of this function should be infinite if we do not assign a domian to them. Hence, I assign $x\in[-3,0]$ here to find the root within this domian. Solving the function by my Newton's method function, I can get a root to be $x=-2.887058$. I draw the function in domain and add the line $f(x)=0$. I also draw the iterations of x's. We can see the last x (the root we got) intersect the line $f(x)=0$ and the function $f(x)$. The codes are in Appendix.
```{r Problem8_sol, eval=T, echo=F, cache=F, include=T, fig.width=6, fig.height=4}
########################### 
#Problem8_sol
###########################
# Solve the function
fun <- function(x){3^x-sin(x)+cos(5*x)}
z <- newton(fun, x0=-3, N=100)
print(z)
curve(3^x-sin(x)+cos(5*x), -3, 0)
abline(h=0)
abline(v=z[1:length(z)], lty=6, col="blue")
plot(1:length(z), z, xlab = "Iterations", ylab = "x's", main = "Iterations on the Path to the Solution")
lines(1:length(z), z)
###########################
```

Repeating this step in different domains and assigning different but resonable initial value of $x_0$, we are able to get many other roots. Since the steps are similar, I omit them in this homework.

## Problem 9
For this problem, we are trying to munge some large datasets and give a briefly summary. Since the datasets are very large, it might be a good idea to read some of them (first 100 rows here) in R and finish the necessary work of columns first. The files are local on my computer. The directory might be different.

```{r Problem9_read, eval=T, echo=F, cache=F, include=T, results='asis'}
    ########################### 
    #Problem9
    #get data  
    ########################### 
    Car_Gebreken_raw <- read.csv("D:/Open_Data_RDW__Gebreken.csv",stringsAsFactors = F, nrows=100, header=T,quote = '"')
    Car_Geconstat_raw <- read.csv("D:/Open_Data_RDW__Geconstateerde_Gebreken.csv", stringsAsFactors = F, nrows=100, header=T)
    Car_Person_raw <- read.csv("D:/Personenauto_basisdata.csv",stringsAsFactors = F, nrows=100, header=T)
    
    Car_Gebreken_raw.colclass <- sapply(Car_Gebreken_raw,class)
    Car_Geconstat_raw.colclass <- sapply(Car_Geconstat_raw,class)
    Car_Person_raw.colclass <- sapply(Car_Person_raw,class)
    
    print("Gebreken")
    print(Car_Gebreken_raw.colclass)

    print("Geconstat")
    print(Car_Geconstat_raw.colclass)

    print("Personen")
    print(Car_Person_raw.colclass)
    ##########################
```

What columns they have are not shown in English which makes me confused. But never mind, I use Google Translate to understand what the columns mean. Continuing with the datasets, I figured that it was unrealistic to read all data in and analyze them because there are so much information in these three datasets. That might be better to focus just one aspect, maybe the data in one year, to be our interest. So I choose rows in 2017 to be my interest just as Dr. Settlage required. And for the columns, I also just choose few of them. For "Gebreken", it has defect code information and description. For "Geconstat", it has inspection date and defect code in respective to license plates. For "Personen", it has make and model of vehicle in respective to license plates. Using Google Translate, I figured which columns I need, first and sixth columns of "Gebreken", first, third, and fifth columns of "Geconstat" (Defects), and first, third, and forth columns of "Person". Thanks to the work from Dr. Settlage, I learned that "fread" is a function which is similar to read.table but much faster and convienient. I use it then to read the specfic data I am interested in.

Next, I use merge function to merge three datasets by plates and defect code. During cleaning data, I translate the columns into English and remove NAs. Then I use unique() function to find out how many different makes and models of cars are reported with defects in 2017. I use sort() function to get which the 5 most frequent defects and the top makes/models having that defects in 2017.

```{r Problem9_munge, eval=T, echo=F, cache=F, include=T, results='asis'}
    ########################### 
    #Problem9
    #munge 
    ########################### 

    # defect code and description
    Car_Gebreken_select <- fread(input = "D:/Open_Data_RDW__Gebreken.csv", header = T, select=c(1,6), showProgress=F)
    # license plate, inspection date and defect code
    Car_Geconstat_select <- fread(input = "D:/Open_Data_RDW__Geconstateerde_Gebreken.csv", header=T, select=c(1,3,5),showProgress=F)
    # license plate, make and model of vehicle
    Car_Person_select <- fread(input = "D:/Personenauto_basisdata.csv", header=T, showProgress = F, select = c(1,3,4))
    
    Car_Geconstat_select_2017 <- Car_Geconstat_select[grep("2017",Car_Geconstat_select$"Meld datum door keuringsinstantie"),]
    # merge datasets
    Car_plates <- merge(Car_Geconstat_select_2017,Car_Person_select,by="Kenteken")
    Car_defect <- merge(Car_Gebreken_select,Car_plates,by="Gebrek identificatie")
    # Clean data
    Car_defect <- Car_defect[complete.cases(Car_defect),]
    # Translate
    colnames(Car_defect) <- c("Defect_Code", "Description", "License_Plate", "Report_Date", "Make", "Model")
    num_makes <- length(unique(Car_defect$Make))
    print(num_makes)
    num_models <- length(unique(Car_defect$Model))
    print(num_models)
    # most frequent defects
    # top5_defects <- sort(table(Car_defect$Defect_Code), decreasing=TRUE)[1:5]
    Top5_defects <- Car_defect %>% 
      group_by(Defect_Code, Make) %>%
      summarise(Make_Count = n()) %>%
      mutate(Defect_Count = sum(Make_Count)) %>%
      arrange(desc(Defect_Count)) %>%
      mutate(Top_Make_Count = max(Make_Count)) %>%
      filter(Make_Count %in% Top_Make_Count) %>%
      select(Defect_Code, Defect_Count, Make, Top_Make_Count)
    knitr::kable(Top5_defects[1:5,],caption="The 5 Most Frequent Defects in 2017")
    # number of defects vs. make
    defect_make <- as.data.frame(table(Car_defect$Make))
    colnames(defect_make) <- c("Make", "Num_Def")
    lm_defect_make <- lm(defect_make$Num_Def~defect_make$Make)
    ano_defect_make <- anova(lm_defect_make, data=defect_make)
    # summary(lm_defect_make)
    knitr::kable(summary(defect_make),caption="Defects Data Summary")
    knitr::kable(anova(lm_defect_make),caption="ANOVA of fitting Number of defects with Make")
    # number of defects vs. model
    defect_model <- as.data.frame(table(Car_defect$Model))
    colnames(defect_model) <- c("Model", "Num_Def")
    # lm_defect_model <- lm(defect_model$Num_Def~defect_model$Model)
    # ano_defect_make <- anova(lm_defect_make, data=defect_make)
    # summary(lm_defect_make)
    # summary(ano_defect_make)
    ########################### 
```

Hence, there are 503 different makes and 33470 different models reported defects in 2017. "AC1 K04 RA2 205 497" are the 5 most frequent defects in 2017. And the regression summary and ANOVA table are shown. For the lm and anova analysis, I finished it for Make but keep getting error "cannot allocate vector of size 8.3 Gb" when doing same thing for Model. I guess that might because of the size of dataframe is to large for lm function. Hope you would not mark me wrong. Besides the output of lm summary is too long, so I do not show it here. The codes are in Appendix.

In order to make the codes work more computationally efficient, I think that might be better to read the data of 2017 only at first.

\newpage
#Appendix 1: R code  
```{r Appendix, ref.label=c("Problem6_func", "Problem6_data", "Problem7_BP_analysis", "Problem8_func", "Problem8_sol", "Problem9_munge"), echo=TRUE, eval=F, tidy=TRUE, include=T}
  
```